An Event-based Architecture for Multi-population Optimization Algorithms

### 2020-01-14 draft.tex
More additions to intro

\begin{abstract} 
Multi-population based methods have been used extensively in
recent years to improve the performance of nature-inspired optimization
algorithms. These methods are parallel and asynchronous by nature and researchers
leverage this to scale implementations 
applying distributed, multi-threaded, parallel, and cloud-based computing. 
% scale implementations with ... implementations. Please rephrase - JJ 
% done - Mario
However, the design  of these cloud-native,  
distributed, multi-population algorithms is not a trivial task. Designers need
to adapt their solutions, 
% design of algorithms is not trivial... designer need to adapt the algorithms... Maybe rephrase to avoid repetition - JJ to leverage
the scalability, elasticity, fault-tolerance, reproducibility, and
cost-effectiveness of cloud systems while conserving the intended functionality.
% I am removing the native part above, because is also very repetetive - Mario 
% Something about "this is the problem we want to address... " to link the previous sentence with this one - JJ 
Furthermore, currently, there are no
cloud-native solutions that support basic algorithm design practices that can make
multi-population methods more efficient: having multiple search strategies and
varying the number and size of subpopulations. We address these problems in this work,
by presenting the design and implementation of a cloud-native 
container-based % again maybe it's better cloud-native instead of reactive - JJ
application for the execution of ensemble multi-population algorithms. Researchers can
use the application to execute multiple experiments, both locally or in a cloud
environment and deploy experiments by specifying the infrastructure as part of
an experiment definition so that results are easily reproducible.
% A sentence on how this implementation is better than the state of the art - JJ 
% Added a comment above - Mario, 
% But maybe we need to say it again? - Mario
% For instance "Generally the current implementations include a single algorithm, which is bad because... " - JJ 
% Also "Current implementations are more or less functional equivalent 
%to parallel, including the fact that they run copies of the same, parametrized, algorithm... 
% If this is too long, maybe just some reference about 
%"current implementation do not leverage the fact that (you can use different algorithms,
% and differently parametrized...) and then qualify this in the introduction - JJ
State of the art approaches use a single-algorithm implementation, and this might
not be the best solution. We compared the speedup, execution time, and the number
of evaluations between a single and an ensemble approach by using Genetic Algorithms (GAs)
and Particle Swarm Optimization (PSO) algorithms against a benchmark for the optimization
of continuous functions. Experiments show that the framework allows the ensemble 
to outperform single-algorithm versions,  
% We also run experiments to see the effect of population size for seveal problem sizes  
% Against NUMBER OF WORKERS (this concept is not mentioned yet)
% We can add something like:      
% Moreover, we evaluated the effect other 
% design parameters had on the performance; in particular, the number and the size of 
% populations concerning the problem size. - Mario
thus proving that cloud-native bioinspired algorithms perform better 
in their "natural" environment than other kinds of algorithms.

%The architecture and the implemented platform is an excellent alternative for
%locally or a cloud-based deployment, 
% thus proving that cloud-native bioinspired algorithms perform better in their "natural" environment that other kind of algorithms - JJ 
\end{abstract}

@@

In the last decades, nature-inspired optimization algorithms have been succesfully
applied to solve complex real-world problems \cite{yang2014nature}. Algorithms
inspired by natural processes include evolutionary algorithms (EAs)
\cite{back1996evolutionary} and swarm intelligence (SI) \cite{kennedy2006swarm},
among others. These population-based algorithms share the common characteristic
of using an initial set of random candidate solutions that are later used to
generate a new set of candidates, using a nature-inspired heuristic. Popular EAs
are Genetic Algorithms (GAs) \cite{holland1992adaptation,eiben2003genetic}, 
Genetic Programming (GP) \cite{back1996evolutionary}, Grey Wolf Optimization
(GWO) and Differential Evolution (DE), while examples of (SI) are particle swarm
optimization (PSO) and ant colony algorithms (ACO).

Population-based algorithms are intrinsically parallel, and can be implemented to run asynchronously. % They needn't be intrinsically asynchronous, I think. Most are implemented synchronouslu - JJ  
%OK - M
The fitness of each individual can be evaluated independently of 
others, and similarly, each population could evolve in isolation. Since 
earlier works, researchers have been proposing some form of parallelization
\cite{muhlenbein1988evolution} to increase the scalability of these algorithms.
The island model was one of the first techniques proposed for parallelization,
which lead to an increased performance \cite{gorges1990explicit,grosso1985computer}. 
The concept was to divide a large population into communicated subpopulations. 
Since then, other population-based algorithms have adopted the concept, 
and researchers have found additional benefits
besides the execution speed; these include avoiding a premature convergence and
maintaining the diversity of the global population. Researchers use the term
multi-population based methods when generally referring to techniques using
subpopulations as part of their optimization strategy. % Some comment about asynchrony would be in order - JJ
% Comments below - Mario
Moreover, parallel implementations can run in two ways: synchronous and
asynchronous. When operations run in parallel but maintain synchronism with
other operations,  all operations need to finish to continue to the next step.
For instance, in a  parent/worker model, the parent needs to wait for all
workers to finish their operations before moving to the next iteration.  In
contrast, in an asynchronous execution,  operations are not synchronized with
each other. In the previous, the parent operation can continue to the next
iteration, even if a worker has not finished its work.  Several works have
proposed asynchronous algorithms \cite{coleman89,baugh2003asynchronous}, reporting benefits in execution time, and
scalability. In other works, sub-populations
are processed independently and communicate asynchronously through a central
repository \cite{sofea:cec2012,JSON} or message queues \cite{salza2019speed, guervos2018introducing} .

When designing efficient multi-population algorithms, researchers need to
consider additional issues. These include the number and size of subpopulations,
the interaction between them, the search area of subpopulations, and the search
strategy and parametrization of each subpopulation. We must take into account
these high-level issues when designing a parallel architecture.

The parallel execution of multi-population algorithms has been the focus of many
works,  going from earlier hardware-based implementations using transputers \cite{gorges1990explicit},
multi-threaded \cite{merelo2019scaling}, and multi-core systems \cite{Serrano2008,lai2019adaptive}. % many citations here - JJ
Recently the trend has been on exploiting a higher number of processing units 
by using GPUs \cite{tan2015survey,li2007efficient}, or distributed systems; techniques used to achieve this include 
web-based \cite{JSON}, map-reduce \cite{fazenda2012},  grid \cite{munawar2010design,Gonzalez09}, 
voluntary computing \cite{MilkyWay}, and cloud computing 
\cite{GValdez2015,salza2019speed,valenzuela2015implementing,FlexGP}. % many references here too - JJ
Cloud computing has become the standard way of running enterprise applications, not
only because of the convenience of the pay-as-you-go model or the non-existent
cost of administration, but also because it offers a way of describing the
infrastructure as part of the code, so that it is much easier to reproduce
results and this has been a boon for scientific computing.  However,  cloud computing
has also been evolving, going from simply putting in the cloud old-style monolithic applications, that is,
applications built on a single stack of services that communicate by layers, to
microservices \cite{microservices} and serverless architectures \cite{varghese2018next, Varghese2018849} that favor the parallel and
asynchronous communication of heterogeneous resources; in the process, 
a new process for designing {\em cloud native applications} has been created, 
with brand new methodologies and techniques \cite{Baldini2016287}. In these architectures,
services or even functions are seen as single processing points,  departing
from the monolithic and even distributed paradigms, become a loosely coupled collection of 
{\em stateless functions} \cite{malawski2017serverless}, that react to events and only come into existence while they 
are doing some processing. % many references more here - JJ

>>>>

\begin{abstract} 
Multi-population based methods have been used extensively in
recent years to improve the performance of nature-inspired optimization
algorithms. These methods are parallel and asynchronous by nature and researchers
leverage this to scale implementations 
applying distributed, multi-threaded, parallel, and cloud-based computing. 
% scale implementations with ... implementations. Please rephrase - JJ 
% done - Mario
However, the design  of these cloud-native,  
distributed, multi-population algorithms is not a trivial task. Designers need
to adapt their solutions, 
% design of algorithms is not trivial... designer need to adapt the algorithms... Maybe rephrase to avoid repetition - JJ to leverage
the scalability, elasticity, fault-tolerance, reproducibility, and
cost-effectiveness of cloud systems while conserving the intended functionality.
% I am removing the native part above, because is also very repetetive - Mario 
% Something about "this is the problem we want to address... " to link the previous sentence with this one - JJ 
Furthermore, currently, there are no
cloud-native solutions that support basic algorithm design practices that can make
multi-population methods more efficient: having multiple search strategies and
varying the number and size of subpopulations. We address these problems in this work,
by presenting the design and implementation of a cloud-native 
container-based % again maybe it's better cloud-native instead of reactive - JJ
application for the execution of ensemble multi-population algorithms. Researchers can
use the application to execute multiple experiments, both locally or in a cloud
environment and deploy experiments by specifying the infrastructure as part of
an experiment definition so that results are easily reproducible.
% A sentence on how this implementation is better than the state of the art - JJ 
% Added a comment above - Mario, 
% But maybe we need to say it again? - Mario
% For instance "Generally the current implementations include a single algorithm, which is bad because... " - JJ 
% Also "Current implementations are more or less functional equivalent 
%to parallel, including the fact that they run copies of the same, parametrized, algorithm... 
% If this is too long, maybe just some reference about 
%"current implementation do not leverage the fact that (you can use different algorithms,
% and differently parametrized...) and then qualify this in the introduction - JJ
State of the art approaches use a single-algorithm implementation, and this might
not be the best solution. We compared the speedup, execution time, and the number
of evaluations between a single and an ensemble approach by using Genetic Algorithms (GAs)
and Particle Swarm Optimization (PSO) algorithms against a benchmark for the optimization
of continuous functions. Experiments show that the framework allows the ensemble 
to outperform single-algorithm versions,  
% We also run experiments to see the effect of population size for seveal problem sizes  
% Against NUMBER OF WORKERS (this concept is not mentioned yet)
% We can add something like:      
Moreover, we evaluated the effect other design parameters had on the
performance; in particular, the number and the size of populations concerning
the problem size. The architecture and the implemented platform is an excellent
alternative for locally or in the cloud, thus proving that cloud-native
bioinspired algorithms perform better in their "natural" environment than other
kinds of algorithms.
\end{abstract}

@@

In the last decades, nature-inspired optimization algorithms have been succesfully
applied to solve complex real-world problems \cite{yang2014nature}. Algorithms
inspired by natural processes include evolutionary algorithms (EAs)
\cite{back1996evolutionary} and swarm intelligence (SI) \cite{kennedy2006swarm},
among others. These population-based algorithms share the common characteristic
of using an initial set of random candidate solutions that are later used to
generate a new set of candidates, using a nature-inspired heuristic. Popular EAs
are Genetic Algorithms (GAs) \cite{holland1992adaptation,eiben2003genetic}, 
Genetic Programming (GP) \cite{back1996evolutionary}, Grey Wolf Optimization
(GWO) \cite{mirjalili2014grey} and Differential Evolution (DE) \cite{karabouga2004simple},
while examples of (SI) \cite{kennedy2006swarm} are particle swarm
optimization (PSO) \cite{clerc2010particle} and ant colony algorithms (ACO) \cite{dorigo1999ant}.

Population-based algorithms are intrinsically parallel, and can be implemented to run asynchronously.
The fitness of each individual can be evaluated independently of 
others, and similarly, each population could evolve in isolation. Since 
earlier works, researchers have been proposing some form of parallelization
\cite{muhlenbein1988evolution} to increase the scalability of these algorithms.
The island model was one of the first techniques proposed for parallelization,
which lead to an increased performance \cite{gorges1990explicit,grosso1985computer}. 
The concept was to divide a large population into communicated subpopulations. 
Since then, other population-based algorithms have adopted the concept, 
and researchers have found additional benefits
besides the execution speed; these include avoiding a premature convergence and
maintaining the diversity of the global population \cite{li2015multi}. Researchers use the term
multi-population based methods when generally referring to techniques using
subpopulations as part of their optimization strategy. % Some comment about asynchrony would be in order - JJ
% Comments below - Mario
Moreover, parallel implementations can run in two ways: synchronous and
asynchronous. When operations run in parallel but must maintain synchronism with
the other operations,  all operations need to finish to continue to the next
step. For instance, in a  parent/worker model, the parent needs to wait for all
workers to finish their operations before moving to the next iteration.  In
contrast, in an asynchronous execution,  operations are not synchronized with
each other. Following the previous example, now the parent operation can
continue to the next iteration, even if a worker has not finished its work.  In
the literature, we can find many asynchronous algorithms
\cite{coleman89,baugh2003asynchronous}, reporting benefits in execution time,
and scalability. In the particular case of asynchronous and cloud-native
multi-population algorithms, recent works propose an asynchronous communication
through a central repository \cite{sofea:cec2012, JSON} or message queues
\cite{salza2019speed, guervos2018introducing}, in this work we follow this
practice.
When designing efficient multi-population algorithms, researchers need to
consider additional issues \cite{Ma2019}. These include the number and size of subpopulations,
the interaction between them, the search area of subpopulations, and the search
strategy and parametrization of each subpopulation. We must take into account
these high-level issues when designing a parallel architecture.

The trend concerning the parallel execution multi-population algorithms  
go from earlier hardware-based implementations using transputers \cite{gorges1990explicit},
multi-threaded \cite{merelo2019scaling}, and multi-core systems \cite{Serrano2008,lai2019adaptive}. % many citations here - JJ
And recently the focus has been on exploiting a higher number of processing units 
by using GPUs \cite{tan2015survey,li2007efficient}, or distributed systems; techniques used to achieve this include 
web-based \cite{JSON}, map-reduce \cite{fazenda2012},  grid \cite{munawar2010design,Gonzalez09}, 
voluntary computing \cite{MilkyWay}, and cloud computing 
\cite{GValdez2015,salza2019speed,valenzuela2015implementing,FlexGP}. % many references here too - JJ
Cloud computing has become the standard way of running enterprise applications, not
only because of the convenience of the pay-as-you-go model or the non-existent
cost of administration, but also because it offers a way of describing the
infrastructure as part of the code, so that it is much easier to reproduce
results and this has been a boon for scientific computing.  However,  cloud computing
has also been evolving, going from simply putting in the cloud old-style monolithic applications, that is,
applications built on a single stack of services that communicate by layers, to
microservices \cite{microservices} and serverless architectures \cite{varghese2018next, Varghese2018849} that favor the parallel and
asynchronous communication of heterogeneous resources; in the process, 
a new process for designing {\em cloud native applications} has been created, 
with brand new methodologies and techniques \cite{Baldini2016287}. In these architectures,
services or even functions are seen as single processing points,  departing
from the monolithic and even distributed paradigms, become a loosely coupled collection of 
{\em stateless functions} \cite{malawski2017serverless}, that react to events and only come into existence while they 
are doing some processing. % many references more here - JJ
